Title of my paper

### 2013-03-22 paper.tex
More equations and nice stuff

\usepackage[utf8]{inputenc}
\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac,wasysym}
\usepackage{listings}
\let\stdsection\section
\renewcommand\section{\newpage\stdsection}

@@

@@

\begin{abstract}
Your abstract goes here...
...
\end{abstract}

@@

@@

@@

@@

lksadjf lakdf jlsak dfjlask dfjlask dfjlsa kdfjasld kfjsald kfjsalkd fjslakd fjlsakd fjlkasd fjlaskd jflsak dfj

@@

\section{Formal Stuff}
10 pagish
This subsection's content...
\subsection{DPL}
Dynamic semantics: http://plato.stanford.edu/entries/dynamic-semantics/

growth of information in time
Inspired by programming language
not philosofically, states can be anything. I don't want to get into that
isomorphic with normal logic (visser siger generalization/projection)
doen't promise to be a perfect dual of natural language, but it's closer, and has been rewarding
Dynamic semantics aims to model meaning and interpretation. You can do that without answering broader philosophical questions
takes the meaning of concepts for given
discourse representation theory, file change, update semantics, 
Hans Kamp, Veltman, Groenendijk

@@

@@

@@

@@

\subsection{Dependencies}
This subsection's content...
\subsection{Montague semantic}
This subsection's content...
\subsection{Semantics via dependencies}
This subsection's content...

@@

\subsection{My implementaiton of DPL semantics}
Maybe this should be in the formal stuff. Nah, there are two versions

Below are some direct implementation of the semantics for DPL. We use the types
\begin{lstlisting}
type Assignment = [(Ref, Val)]
type Relation = [Assignment] -> [Assignment]
\end{lstlisting}
This differs from the indirect set of types are used in the earlier formalisms:
\begin{lstlisting}
type Assignment = [(Ref, Val)]
type Relation = [Assignment] -> [Assignment] -> Bool
\end{lstlisting}
The later is very difficult to implement, since the semantics require quantification over all assignments. Like for composition.

@@

@@

\section{Perspectives}
2pagish
What I learned

\section*{Acknowledgements}
Thanks to Samson
This section's content...\cite{visser1999donkey}
the PCFG parser\cite{klein2003accurate}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac,wasysym,listings}
\usepackage{tikz}
\usepackage{enumitem}

\usetikzlibrary{shapes,arrows}

\let\stdsection\section
\renewcommand\section{\newpage\stdsection}

@@

\lstloadlanguages{Haskell}
\lstnewenvironment{haskell}
    {\lstset{}%
      \csname lst@SetFirstLabel\endcsname}
    {\csname lst@SaveFirstLabel\endcsname}
    \lstset{
      basicstyle=\small\ttfamily,
      escapeinside={/+}{+/},
      flexiblecolumns=false,
      xleftmargin=0.05\textwidth,
      basewidth={0.5em,0.45em},
      literate={+}{{$+$}}1
               {/}{{$/$}}1
               {*}{{$*$}}1
               {=}{{$=$}}1
               {>}{{$>$}}1
               {<}{{$<$}}1
               {\\}{{$\lambda$}}1
               {++}{{$+\!\!\!+$}}1
               {::}{{$:\!\!\!:$}}1
               {\\\\}{{\char`\\\char`\\}}1
               {->}{{$\rightarrow$}}2
               {>=}{{$\geq$}}2
               {<-}{{$\leftarrow$}}2
               {<=}{{$\leq$}}2
               {=>}{{$\Rightarrow$}}2
               {/=}{{$\neq$}}2
               {.}{{$\circ$}}2
               {<<}{{$\ll$}}2
               {>>}{{$\gg$}}2
               {>>=}{{$\gg\!=$}}2
               {>=>}{{$>\!=\!>$}}2
               {|}{{$\mid$}}1
               {undefined}{{$\bot$}}1
               {`elem`}{{$\in$}}1
    }

@@

\begin{abstract}
Your abstract goes here...

Érica is super hot

‘a brief description of what you did; no more than a page.’
The abstract provides a summary of the project report. It may contain some or all of the following moves, 
often, but not always, in the following order. Only use one or two sentences for each move. Keep the 
language and sentence structure simple. The reader should be able to read the abstract and obtain the 
necessary information quickly and efficiently.
Move 1:  Background to the project
Move 2: Purpose of the project
Move 3: Problem tackled
Move 4: Work carried out
Move 5: Results 
Move 6: Conclusions or implications
Move 7: Achievements of the project

\end{abstract}

@@

Move 1 Background 
Possible Steps 1. Why the area is important
2. Giving background information
3. Reviewing previous research
Move 2 Indicating a Problem or Need
Move 3 Presenting the Project
Possible Steps 1. Purposes, aims, objectives
2. Work carried out
3. Justification or importance of the project
4. Outline of the structure of the report

@@

% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20,
    text width=4.5em, text badly centered, node distance=2.5cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20,
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']

\begin{tikzpicture}[node distance = 3cm, auto]
    % Place nodes
    \node [block] (nl) {Natural Language};
    \node [block, right of=nl] (dr) {Dependency Representation};
    \node [block, right of=dr] (monad) {Enhanced DPL Representation};
    \node [block, right of=monad] (dpl) {Dynamic Propositional Logic};
    \node [block, right of=dpl] (prop) {First Order Logic};
    % Draw edges
    \path [line] (nl) -- (dr);
    \path [line] (dr) -- (monad);
    \path [line] (monad) -- (dpl);
    \path [line] [dotted] (dpl) -- (prop);
\end{tikzpicture}

@@

Additionally, the indefinite article `a' is normally understood as an existential quantifier

Week and strong readings, check this out \cite{kanazawa1994weak}

@@

This one has problems with scope
\begin{equation}
\forall x (farmer(x) \wedge \exists y (donkey(y) \wedge owns(x,y) \rightarrow beats(x,y)))
\end{equation}

Moving the scope doesn't solve it, since now if there are two donkeys and he only owns one of them, the sentence will be true nomatter if he beats them or not. Also if he owns say a donkey and a horse, the left part of the implication would be vacously true and hence the beats part wouldn't matter.
\begin{equation}
\forall x \exists y (farmer(x) \wedge donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
\end{equation}

@@

More stuff from wikipedia:

Unfortunately, this translation leads to a serious problem of inconsistency. Indefinites must sometimes be interpreted as existential quantifiers, and other times as universal quantifiers, without any apparent regularity

More questions:

Should those three have the same meaning?
[1a]  Jane likes a painting.
[1b]  It is not the case that Jane doesn't like a painting.
[1c]  Jane likes a painting and either it is raining or it isn't.

\section{Formal Stuff}
10 pagish

This subsection's content...

Background should introduce any information that is necessary for the examiner to understand your project. 
The information here is generally more specific than the background given in the Introduction.
Requirements gives the program requirements. These chapters prepare the ground for the Design chapter.
That is: What do we need to handle to be able to solve this problem

In particular, anaphora resolution is part of DRT, but in dynamic semantic systems anaphora resolution is assumed to be performed by a separate component of the theory

\subsection{DPL}
Dynamic semantics: http://plato.stanford.edu/entries/dynamic-semantics/

growth of information in time
Inspired by programming language
not philosofically, states can be anything. I don't want to get into that
isomorphic with normal logic (visser siger generalization/projection)
doen't promise to be a perfect dual of natural language, but it's closer, and has been rewarding
Dynamic semantics aims to model meaning and interpretation. You can do that without answering broader philosophical questions
takes the meaning of concepts for given
dynamic assignment language, discourse representation theory, file change, update semantics, 
Jan van Eijck, Hans Kamp, Veltman, Groenendijk

@@

hvad med sudoku. I started er alle assignments mulige. Nae.. Men. hm

@@

\subsection{Connection with first order logic}

normal logik kan også laves om til dpl. Nemt, bare brug not.not til at scope E.

An alternative way of defining the semantics

Shows that we aren't doing anything `odd' or `fancy'

shows sound and complete wrt.

Preconditions
weakest phi sa <phi> pi <T> is still true

Like hoare stuff

inspired Dynamic assignment language \cite{eijck1992dynamic}
Jan van Eijck 
Fer-Jan de Vries

like using a stack

$\alpha \models \langle\psi\rangle\phi$ iff there is an assignment $\beta$ with $\alpha[\psi]\beta$, and $\beta \models \phi$

Definition of true and false $\alpha\models\phi :\leftrightarrow \exists\beta(alpha\phi\beta)$.

%{

%Det her skal med for at vi kan forsvare ligningerne

%α[⊥]β :⇔ 0 ≠ 0.
%α[⊤]β :⇔ α = β.
%α[P(x0, …, xn−1)]β :⇔ α = β and ⟨αx0, …, αxn−1 ⟩ ∈ I(P), 
%where P is a predicate symbol of Σ with arity n.
%α[∃v]β :⇔ α[v]β, where α[v]β iff αw = βw, for all variables w ≢ v.
%α[φ · ψ]β :⇔ ∃γ  α[φ]γ[ψ]β.
%α[~(φ)]β :⇔ α = β and ∀γ  ¬ α[φ]γ.
%

Det giver ikke mening at skrive $\not\models[\psi]$ da typen af $[\psi]$ er en relation. Eller hvad?

\begin{align}
\alpha[\psi_1\cdot\psi_2]\beta &:\leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \label{sem_and} \\
\alpha[\psi_1\cup\psi_2]\beta &:\leftrightarrow \alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta \label{sem_or} \\
\alpha[P(x_1,\dots,x_n)]\beta &:\leftrightarrow \alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \label{sem_pred} \\
\alpha[\neg\psi]\beta &:\leftrightarrow \alpha = \beta \wedge \alpha\not\models[\psi] \label{sem_neg}\\
\alpha[\exists x]\beta &:\leftrightarrow \forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \label{sem_exists} \\
\intertext{Using these we can derive the following:}
\alpha[\bot]\beta &\leftrightarrow \alpha = \beta \wedge \bot \nonumber\\
                  &\leftrightarrow \bot \label{sem_bot}\\
\alpha[\top]\beta &\leftrightarrow \alpha = \beta \wedge \top \nonumber\\
                  &\leftrightarrow \alpha = \beta \label{sem_top}\\
\intertext{Taking $\psi_1\rightarrow\psi_2$ to mean $\neg(\psi_1\wedge\neg\psi_2)$ we also get the following neat equation, originally by Groenendijk and Stokhof\cite{groenendijk1991dynamic}:}
\alpha[\psi_1\rightarrow\psi_2]\beta
 &:\leftrightarrow \alpha[\neg(\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
 &\leftrightarrow \alpha = \beta \wedge \alpha\not\models[\psi_1\wedge\neg\psi_2] \nonumber\\
 &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\alpha[\psi_1\wedge\neg\psi_2]\gamma) \nonumber\\
 &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta[\neg\psi_2]\gamma)) \nonumber\\
 &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta=\gamma\wedge\delta\not\models[\psi_2])) \nonumber\\
 &\leftrightarrow \alpha = \beta \wedge \forall\gamma(\forall\delta(\delta=\gamma\rightarrow(\alpha[\psi_1]\delta\rightarrow\delta\models[\psi_2]))) \nonumber\\
 &\leftrightarrow \alpha = \beta \wedge \forall\gamma(\alpha[\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \label{sem_impl}
\end{align}

Hence we can read implication as a the filter, that if you can get `through' $\psi_1$ then you must also be able to get through $\psi_2$. Also notice that $\gamma$ doesn't have to equal $\alpha$. Hence variables introduced in $\psi_1$ can bind variables in $\psi_2$. This will show to be extreamly important in the treatment of donkey sentences later.

Talking about binding, also notice that our union or `or' construct is a bit special. $\psi_1$ and $\psi_2$ cannot bind variables in each other, but in $(\psi_1\cup\psi_2)\cdot\psi_3$ they can both bind variables in $\psi_3$. This is going to introduce a lot of weird situations later. An alternative definition of `or': $\neg(\neg\psi_1\cdot\neg\psi_2)$ will also be considered.

For \eqref{sem_pred} $P$ is a predicate symbol of $\Sigma$ with arity n.

static negation. DOesn't allow new things. LIke a filter

some people have a closed/static or: $\neg(\neg\psi_1\wedge\neg\psi_2)$ \cite{groenendijk1991dynamic}

%Notice that \eqref{sem_bot} and \eqref{sem_top} are special cases of \eqref{sem_pred}.

We are going to introduce formuals that allow us to recursively convert any DRA formula into PL.

To smoothen the following proofs, we will introduce a special syntax `$\langle\psi\rangle\phi$' where $\psi$ is a DRA formula, $\phi$ is a PL formula and the whole thing is a PL formula. For some assignment $\alpha$, we define `$\alpha\models\langle\psi\rangle\phi$' to mean $\exists\beta(\alpha[\psi]\beta \wedge \beta\models\phi)$.

\begin{align}
\alpha\models\langle\exists x\rangle\phi
 & \leftrightarrow \exists\beta (\alpha[\exists x]\beta \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta (\forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta (\exists\nu (\forall\omega (\alpha[x:=\nu]_{\omega} = \beta_{\omega})) \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta (\exists\nu (\alpha[x:=\nu] = \beta) \wedge \beta\models\phi) \nonumber\\
% & \leftrightarrow \exists\nu (\exists\beta (\alpha[x:=\nu] = \beta \wedge \beta\models\phi)) \nonumber\\
 & \leftrightarrow \exists\nu (\alpha[x:=\nu]\models\phi) \nonumber\\
 & \leftrightarrow \alpha\models\exists x (\phi) \label{conv_exists}\\\nonumber\\
%
\alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
 & \leftrightarrow \exists\beta (\alpha[\psi_1\cdot\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta(\exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \exists\beta(\gamma[\psi_2]\beta \wedge \beta\models\phi)) \nonumber\\
 & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
 & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi \label{conv_and}\\\nonumber\\
%
\alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
 & \leftrightarrow \exists\beta (\alpha[\psi_1 \cup \psi_2]\beta \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta ((\alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta (\alpha[\psi_1]\beta \wedge \beta\models\phi) \vee \exists\beta (\alpha[\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow (\alpha\models\langle\psi_1\rangle\phi) \vee (\alpha\models\langle\psi_2\rangle\phi) \nonumber\\
 & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\\nonumber\\
%
\alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
 & \leftrightarrow \exists\beta (\alpha[P(x_1,\dots,x_n)]\beta \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta (\alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \alpha\models\phi \nonumber\\
 & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi \label{conv_pred}\\\nonumber\\
%
\alpha\models\langle\neg\psi\rangle\phi
 & \leftrightarrow \exists\beta (\alpha[\neg\psi]\beta \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \exists\beta (\alpha=\beta \wedge \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \beta\models\phi) \nonumber\\
 & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \alpha\models\phi \nonumber\\
 & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma \wedge \gamma\models\top) \wedge \alpha\models\phi \nonumber\\
 & \leftrightarrow \alpha\not\models\langle\psi\rangle\top \wedge \alpha\models\phi \nonumber\\
 & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi \label{conv_neg}
\end{align}

Notice that $\alpha\models\langle\bot\rangle\phi$ and $\alpha\models\langle\top\rangle\phi$ follow trivially from \eqref{conv_pred}.

\begin{haskell}
addDpl :: Prop -> [DPL] -> Prop
addDpl p [] = p
addDpl p (DF:ds) = F
addDpl p (DT:ds) = addDpl p ds
addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
addDpl p ((DExists k):ds) = addDpl (Exist k p) ds
addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards
addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
\end{haskell}

@@

We can also try to define semantics using first order logic. This totally doesn't work and convert reasonable statements into mess. In particular it is hard to have or not being closed scopewise.

The difference between those two:

$addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds$

$addDpl p ((DOr d1 d2):ds) = Or (addDpl p (d1:ds)) (addDpl p (d2:ds))$

Fra Dynamic Montague Grammar:
This means that the extended dynamic semantics is also able to deal with a donkey
disjunction such as (49), and in such a way that it is equivalent with the donkey
conditional (50). It can further be noted that this notion of dynamic disjunction is
like dynamic implication in this respect that sentences which follow it, are in fact
conjoined with its second argument:
Fact 24 [either p or q] ; r = either p or [q ; r]
This is needed for a treatment of examples such as:
(54) Either there is no bathroom here, or it is in a funny place. In any case, it is
not on the ground floor.

Den ene giver mulighed for at binde fra a til b. Den anden giver mulighed for at begge dele binder res

\subsubsection{Weird disjunctive subjects}

problem wouldn't work in Spanish
A farmer or their donkey was arrested this morning.

\begin{equation}
Ex(farmer(x)) ^ Ey(donkey(y)) ^ (a-t-m(x) v a-t-m(y))
ExEy(farmer(x) ^ donkey(y) ^ owns(x,y) ^ (a-t-m(x) v a-t-m(y)))
Ex(farmer(x)) v Ey(donkey(y))
((Ex . farmer(x) . Es.s=x) v (Ey . donkey(y) . owns(x,y) . Es.s=y)) . a-t-m(s)
((Ex . farmer(x) . (Es.s=x v (Ey . donkey(y) . owns(x,y) . Es.s=y)) v (Ey . donkey(y) . owns(x,y) . Es.s=y)) . a-t-m(s)
\end{equation}

If a farmer is Spanish or a Japanese she beats her donkey
It is a donkey or a horse
You beat it or I do
Either a farmer or his donkey is happy

either = xor?

Vissers dynamic bracketing

\subsection{Errors}

Or should this be in implementation? Hard for me to say

@@

\subsection{Montague semantic}
Montague grammars are an approach to combinatorical semantics based on types and lambda calculus. It is interesting because it 
Types
Quantification

montague can make any logic compositional

Types of:
  One-place predicate constant (farmer, sleeps, walks): Entity -> Bool
  Transitive verb (owns, beats): Entity -> Entity -> Bool
  Attributive adjective (good, inteligent, former): (Entity -> Bool) -> (Entity -> Bool)
  Noun phrase: (Entity -> Bool) -> Bool
  Proper noun: Entity
  Quantifiers/Determiners: a, the, every, no: (Entity -> Bool) -> (Entity -> Bool) -> Bool
  Sentence: Bool

\subsection{Weak and strong readings}

Let's return multiple intepretations

drt

The upshot of the foregoing observations is that, apparently, indefinites are neither quantifiers nor referential terms, and this problem entrains another one, for as long as it unclear what indefinites mean, it will also remain obscure how they can serve as antecedents to pronouns

Is this the same as the problem of multiple `scope bearing operators?'
Sentences with multiple scope bearing operators - e.g.,
quantified noun phrases or negations - are often ambiguous.

%$• “Every student reads a book”
%- ∀x(student(x) ⇒ ∃y(book(y) ∧ read(y)(x)))
%- ∃y(book(y) ∧ ∀x(student(x) ⇒ read(y)(x)))

\subsection{Dependencies}

Extracted from parse trees using tregex\cite{de2006generating}

\tikzstyle{level 1}=[level distance=2.5cm, sibling distance=3.5cm]
\begin{figure}
\centering
\begin{tikzpicture}[grow=down, sloped]
\node {root}
    child {
        node {owns}
        child {
            node {If}
            edge from parent
            node[above] {mark}
        }
        child {
            node {farmer}
            child {
                node {a}
                edge from parent
                node[above] {det}
            }
            edge from parent
            node[above] {nsubj}
        }
        child {
            node {beats}
            child {
                node {he}
                edge from parent
                node[above] {nsubj}
            }
            child {
                node {it}
                edge from parent
                node[above] {dobj}
            }
            edge from parent
            node[above] {advcl}
        }
        child {
            node {donkey}
            child {
                node {a}
                edge from parent
                node[above] {det}
            }
            edge from parent
            node[above] {dobj}
        }
        edge from parent
        node[above] {root}
    };
\end{tikzpicture}
\caption{M1} \label{fig:M1}
\end{figure}

\begin{equation}
NP < NN \$ VP
\end{equation}

An NP over an NN and w/ sister VP

\begin{equation}
NP < (NN < dog) \$ (VP <<\# (barks > VBZ))
\end{equation}

An NP both over an NN over `dog' and with
a sister VP headed by `barks' under VBZ

\subsection{Corefferences}
Corefferences (or anaphora) is essential in order to create advanced discourse, since it is what allows us maintain a topic beond the first mention. Without corefferences we couldn't even express simple logical statements such as transitivity: `a man's brother's brother is his brother' etc.

Identifying corefferences (or resolving anaphora) is hence an important part of converting natural language into logic. The issue is not touched upon in Visser's paper, which casually sidesteps the problem by assigning variables to all mentioned entities.

For the subtask of resolving simple pronouns a solution could be to work in terms of subjects, objects, indirect objects and so on. However if not earlier, this certainly fails for pronouns seperated from their antecedent by sentence bounderies. Also we want to resolve other kinds of corefferences, such as `John, the farmer, owns a donkey', where `the farmer' and `John' reffers to the same entity.

It quickly becomes clear that the problem is not only syntactical. In the sentence `John and his wife had a farm. He took care of the donkeys' we could substitute `He' with `She' and to change the refferant of the pronoun. In this case we need semantic knowledge of genders to proceed.

In my project I have taken advantage of the corefference tagger built into Stanford CoreNLP\cite{lee2013deterministic}\cite{lee2011stanford}\cite{raghunathan2010multi}. In the following section I will summerize the workings of this machine, made to work across large texts as well as simple sentences.

Correference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large itterative applications of linguistic heuristics. Of course the procedure assumes (most likely) machine learned postagging and named entity recognization, but the main algorithm is transparent and deterministic.

The idea is to iteratively go over the text with different `sieves'. The below example explains each step with an example I have adopted from \cite{lee2013deterministic}:

\begin{quotation}
John is a farmer. He owned a stubborn donkey.
A girl was looking at the donkey.
``It was my favorite,'' John said to her.
\end{quotation}

\begin{description}
\item[Mention detection]
The first step is to identify all noun phrases (NP) and personal (PRP) and possesive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:

[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
``[It]$_7^7$ was [[my]$_9^9$ favorite]$_8^8$,'' [John]$_{10}^{10}$ said to [her]$_{11}^{11}$.

All mentions are initially assigned to seperate entities (entity ids). The entities are also attached information we can easily pull from the postree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
\item[Speaker Sieve]
The purpose of this first sieve is to link self referring pronouns to speakers. Speakers are identified simply by their connection to verbs such as `say' and proximity to the quotations. In our case the pronoun `my' gets linked to `John':

[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
``[It]$_7^7$ was [\textbf{[my]$_9^1$} favorite]$_8^8$,'' \textbf{[John]$_{10}^{9}$} said to [her]$_{11}^{11}$.

In conversational text we would store information about the speakers in our entities. Thus we could e.g. match mentions of `you' to the speaker of the previous quote we saw.

We also note restrictions for future use, so we don't risk corefferencing an `I' with a `he' etc.
\item[String Match]
The second step is to merge entities reffered to by exactly the same string. In our case we have two mentions of `John':

\textbf{[John]$_1^1$} is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^8$,'' \textbf{[John]$_{10}^{1}$} said to [her]$_{11}^{11}$.
\item[Relaxed String Match]
This third sieve doesn't get activated by our example. It will try to remove relative clauses and other text following the head word of a mention, and do an exact match of the result. For example [John] and [John, whose donkey loves him] will correctly be merged.

Like the `String Match' sieve, we are not doing anything fancy here, and we will incorrectly corefference the two mentions of `John' in: `[John, who loves donkeys] and [John, who hates donkeys] were brothers.'
\item[Precise Constructs]
The `Precise Constructs' sieve uses a lot of common patterns that link metions. In our case two `X is Y' patterns are found:

\textbf{[John]$_1^1$} is \textbf{[a farmer]$_2^1$}. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
``\textbf{[It]$_7^7$} was \textbf{[}[my]$_9^1$ \textbf{favorite]$_8^7$},'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.

Other patterns include acronyms and appositives. For example [Canadian Donkey \& Mule Association] and [CDMA] are linked, because the second is tagged as an acronym and it matches the upper case letters of the first one. An appositive is usually a pattern `X, Y, ...' where Y is another description of X.
\item[Strict Head Match A, B, C] are simply rules strip mentions down to their head word and try to find exact matches. This is similar to `Relaxed String Match', but more radical. In our case we correctly get a link between `a stubborn donkey' and `the donkey':

[John]$_1^1$ is [a farmer]$_2^1$. [He]$_3^3$ owned \textbf{[a stubborn donkey]$_4^4$}.\newline
[A girl]$_5^5$ was looking at \textbf{[the donkey]$_6^4$}.\newline
``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^7$,'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.

Linking mentions based on their head word is dangerous. e.g. `University of Oxford' and `University of Cambridge' both have `University' as their head word, but are (obviously?) different entities. To combat this a lot of strict rules must be observed. These are however gradually weakened in sieve B and C.
\item[Proper Head Noun Match] could be called `Strict Head Match D'. It is the weakest form of the sieve, as it has only the three contraints: two mentions of the same entity cannot have one included in the other, they cannot contain words reffering to conflicting geographical locations, and they cannot have different numbers. e.g. [donkeys] and [at least 5 donkeys].
\item[Pronoun Match]
This last sieve is perhaps the most important one. It is done in the end so that we may have as much data in our entities as possible for matching. Using guesses on gender, number and animacy we can make our final merges:

\textbf{[John]$_1^1$} is [a farmer]$_2^1$. \textbf{[He]$_3^1$} owned [a stubborn donkey]$_4^4$.\newline
\textbf{[A girl]$_5^5$} was looking at \textbf{[the donkey]$_6^4$}.\newline
``\textbf{[It]$_7^4$} was [[my]$_9^1$ favorite]$_8^4$,'' [John]$_{10}^1$ said to \textbf{[her]$_{11}^{5}$}.

As always there are a few contraints put on what can be merged. One rule is that the distance between pronun and antecedent must be maximum 3 sentences.
\end{description}

After the final step different post processing options are usually performed. For example singleton mentions are often discarded. That would mean we got rid of `a musician' and `my favorite'.

The reason the rules are applied in the order above is to give the most certain rules the highest priority.

The Stanford procedure above is currently the strongest competitor in various correference competitions.

\subsection{Semantics via dependencies}
This subsection's content...

@@

There is not a big first order logic component in my project, but in the following it will often be useful to refer to the following structure

\begin{haskell}
data Prop =
    T | F
  | Not   Prop
  | And   Prop Prop
  | Or    Prop Prop
  | Predi Ref [Ref]
  | Exist Ref Prop
  | All Ref Prop
  | Impl Prop Prop
\end{haskell}

And since we are going to deal with some large computer generated formulas, I've implemented a simple simplification function, which simlply applies basic identities for $\top$ and $\bot$, and a comnbination of De Morgan's laws and the implication identity to eliminate negations.

One interesting part is that since the simplifyer has no look ahead, it will sometimes not find all applicable simplifications in one pass. For example
%
\begin{haskell}
simplify (Not (Impl p F)) = Not (simplify (Impl p F))
                          = Not (Not simplify p)
                          = Not (Not q)
\end{haskell}
%
Hence we need to run it a couple of times to get a good result. We can run it until the output doesn't change, or even simpler, just use a function like
%
\begin{haskell}
fexp :: (a -> a) -> Int -> (a -> a)
fexp f e = (iterate (f.) id) !! e
\end{haskell}

\subsection{My implementaiton of DPL semantics}

Syntactically we will be using the following structure

\begin{lstlisting}
data DPL =
    DT | DF
  | DComp DPL DPL
  | DOr DPL DPL
  | DNot DPL
  | DImpl DPL DPL
  | DExists Ref
  | DPredi Ref [Ref]
\end{lstlisting}

Below are some direct implementation of the semantics for DPL. We use the types
\begin{lstlisting}
type Assignment = [(Ref, Val)]
type Relation = Assignment -> [Assignment]
\end{lstlisting}
This differs from the indirect set of types are used in the earlier formalisms:
\begin{lstlisting}
type Assignment = [(Ref, Val)]
type Relation = Assignment -> Assignment -> Bool
\end{lstlisting}

intuitive

uses well known isomorphism

more direct

The later is very difficult to implement, since the semantics require quantification over all assignments. Like for composition.

@@

\subsection{Handling variables}
One thing I haven't found much \cite{visser1999donkey}
Stanford: \cite{lee2013deterministic}
\cite{lee2011stanford}
\cite{raghunathan2010multi}
Coreferences
Using the parse tree and NPs

Something that's very interesting is when to use free variables and when to use bound variables. We can think of free variables as actors known by the context and bound variables as actors introduced in the sentence. So if we had `The farmer had a friend. His name was Anders', we could write the first part as $\exists y \cdot friends(x,y)$ and the second part as $name(y, `Anders')$. This is important to keep compositionality, but it does of course raise the interesting question of how to make sure that the variable for `Anders' in the first and the second sentence is the same.

In addition to being essential for composition, free variables are also very useful for information extraction type sentences. We can immagine writing a curious researcher writing into a search engine `farmers who like donkeys' which translates into $farmer(x) \cdot likes(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y)$.

So how do we know when to include an $\exists$ and when not to?

The coreference parser does however not find entities that are only mentioned once. Hence to make the later processing simpler, we assign those entitie to free variable now as well.

The task is slightly complicated by the fact that NP's can be nested, so `The farmer and the donkey' is a NP consisting of two smaller NP's `The farmer' and `the donkey' seperated by `and'.

We resolve this by assigning top down, overwriting `word to ref' assignments with fresh variables as we go deeper down the tree. This way, for our short sentence, we might end up with the assignment $\{`The': `y', `farmer': `y', `and': `x', `the': `z', `donkey': `z'\}$

I parse the parse tree using parsec\cite{leijen2001parsec}

\begin{lstlisting}
toMapping :: [Ref] -> PosTree Word -> ([Ref], Map Int Ref)
toMapping vs (Leaf _ _) = (vs, empty)

toMapping vs tr@(Phrase "NP" subs) = (vs', submaps `union` np)
  where np = mapInterval 0 (tsize tr) v
        (v:vs', submaps) = toMapping vs (Phrase "" subs)

toMapping vs (Phrase _ []) = (vs, empty)
toMapping vs (Phrase _ (sub:subs)) = (vs'', p `union` shifted)
  where (vs', subm) = toMapping vs (Phrase "" subs)
        shifted = mapKeys (+tsize sub) subm
        (vs'', p) = toMapping vs' sub
\end{lstlisting}

where $tsize :: PosTree Word -> Int$ gives the number of words in the tree

@@

\subsection{Examples}
Until now we've been very focused on the same one or two simple sentences. To provide a better test bench for our approach, it is interesting to follow some sentences through the different levels of our approach.
